### 2024-10-22 manuscript.tex
Fix grammar

Increasing climate model resolution is a common way of improving model accuracy
\citep{mauritsen2022}. This is of course limited by the available computational
power and a tradeoff with increasing parametrisation complexity. Current
computational availability and acceleration from general-purpose computing on
graphics processing units (GPGPUs) is progressing to enable km-scale (also
called k-scale) Earth system models (ESMs) and coupled atmosphere--ocean
general circulation models (AOGCMs) in research conditions today and
operationally in the forthcoming years. Therefore, it represents a natural
advancement in climate modeling.  Global strom-resolving models (GSRMs) are
emerging as a new front in the development of high-resolution global climate
models, with horizontal grid resolutions of about 2--8 km
\citep{satoh2019,stevens2019}. This is enough to resolve mesoscale convective
storms, but smaller-scale convective plumes and smaller-scale cloud structure
remain unresolved. At a 5-km scale, non-hydrostatic processes also become
important, and for this reason such models are generally non-hydrostatic. The
terms global cloud-resolving models, or global convection-permitting/-resolving
models are also sometimes used interchangeably with GSRMs but imply that clouds
or convection are resolved explicitly, which is not entirely true for GSRMs, as
this would require an even higher horizontal resolution \citep{satoh2019}.
Representative of these efforts is the DYnamics of the Atmospheric general
circulation Modeled On Non-hydrostatic Domains (DYAMOND) project
\citep{stevens2019,dyamond}, which is an intercomparison of nine global
storm-resolving models (GSRMs) on two 40-day time periods in summer (1 August
-- 10 September 2016) and winter (20 January -- 1 March 2020). A new one-year
GSRM intercomparison is currently proposed by \cite{takasuka2024}, with the
hope of also evaluating the seasonal cycle and large-scale circulation.  An
alternative to using a computantionally costly GSRM is to train an artificial
neural network on GSRM output and use it for subgrid-scale clouds, as done with
the GSRM ICON by \cite{grundner2022} and \cite{grundner2023}.

@@

The Southern Ocean (SO) is known to be a problematic region for climate model
biases due to reasons such as a lack of surface and in situ observations and
being a lower priority region for numerical weather prediction (NWP) and
climate model development because it is far away from populated areas of
interest.  Nevertheless, radiation biases and changes over an area of its size
have a very substantial influence on the global climate, and the SO is an
important part of the global ocean conveyor belt.  Marine clouds have
disproportionate effect on top of atmosphere (TOA) SW radiation due to the
relatively low albedo of the sea surface.  The relative longitudinal symmetry
of the SO means that model cloud biases tend to be similar across longitudes.
Here, we conventionally refer to the SO as ocean regions south of 40°S,
low-latitude SO as 40--55°S and high-latitude SO as south of 55°S.

@@

The main aim of this study is to evaluate the GSRM version of ICON, developed
jointly by nextGEMS, Deutscher Wetterdienst, Max-Planck-Institute for
Meteorology, Deutsches Klimarechenzentrum (DKRZ), Karlsruhe Institute of
Technology, and the Center for Climate Systems Modeling.  Previous studies have
identified substantial large-scale biases in climate model clouds over the SO,
affecting sea surface temperature and the Earth’s albedo. Our aim is to
quantify how well the GSRM ICON is simulating clouds in this region,
particularly in light of the fact that subgrid-scale clouds and convection are
not parameterised in this model. This region is mostly dominated by boundary
layer clouds generated by shallow convection, and these are problematic to
observe by spaceborne lidars and radars, which are affected by attenuation by
overlapping and thick clouds and ground clutter, respectively.  Specifically,
the radar on CloudSat and lidar on CALIPSO (neither of them operational any
more) are affected by the abovementioned issues, resulting in a strong
underestimation of cloud occurrence below 2 km relative to ground-based lidar
observations \citep{mcerlich2021}.  This, in turn, can lead to systematic
biases in low clouds in climate models, which are frequently evaluated against
CloudSat--CALIPSO products. Reanalyses can also suffer from cloud biases, as
these are usually parametrised in their atmospheric component, and also in
regions where input observations are sparse.  This makes them a problematic
reference for clouds over the SO, and any biases relative to a reanalysis
should be interpreted with caution. Instead, we chose to use a large set of
ship-based observations conducted with ceilometers and lidars on board of the
RV \emph{Polarstern} and other voyages and stations as a reference for the
model evaluation.

Altogether, we analysed about 2400 days of data from 31 voyages and one
sub-Antarctic station covering diverse longitudes and latitudes of the SO. To
achieve a like-for-like comparison with the model, we used a ground-based lidar
simulator called the Automatic Lidar and Ceilometer Framework [ALCF;
\cite{kuma2021}]. We contrasted the results with the European Centre for
Medium-Range Weather Forecasts (ECMWF) Reanalysis 5 [ERA5; \cite{era5}] and the
Modern-Era Retrospective analysis for Research and Applications, Version 2
[MERRA-2; \cite{gelaro2017}].

@@

Together, we analysed data from 31 voyages of RV \emph{Polarstern}, resupply
vessel (RSV) \emph{Aurora Australis}, RV \emph{Tangaroa}, RV \emph{Nathaniel B.
Palmer}, Her Majesty's New Zealand Ship (HMNZS) \emph{Wellington} and one
sub-Antarctic station (Macquarie Island) in the SO south of 40°S between 2010
and 2021. Fig. \ref{fig:map} shows a map of the voyages, Table
\ref{tab:voyages} list the voyages, campaigns and stations, and Table
\ref{tab:voyage-references} lists references where available. Altogether, the
voyages and station dataset comprised 2421 days of data south of 40°S, but the
availability of ceilometer data was slightly smaller due to gaps in
measurements.

@@

The data sources contained ceilometer observations captured by the Vaisala CL51
 perating at a wavelenght of 910 nm, Vaisala CT25K operating at 905 nm, and
Lufft CHM 15k operating at 1064 nm, described in detail below (Sections
\ref{sec:cl51} and \ref{sec:chm15k}). A ceilometer is a low-power near-infrared
vertically pointing lidar with the purpose of measuring the cloud base, but it
also measures the full vertical structure of clouds as long as the laser signal
is not attenuated by thick clouds, which can be used to infer additional
information such as a cloud mask and cloud occurrence by height.

@@

\begin{table}[p!]
% TODO: Add sampling intervals.
\caption{
An overview of the analysed voyages, campaigns and stations. Start, end, and
the number of days (UTC; inclusive) refer to the time period when the vessel
was south of 40°S.  Abbreviations: ceilometer (ceil.), Australia (AU), New
Zealand (NZ), South America (SA), Atlantic Ocean (AO), and Africa (AF). The
number of days is rounded to the nearest integer. CL51/31 indicates CL51
configured to emulate CL31.
}
\label{tab:voyages}
\small
\begin{tabular}{llllllr}
\textbf{Name} & \textbf{Vessel or station} & \textbf{Ceil.} & \textbf{Region} & \textbf{Start} & \textbf{End} & \textbf{Days}\\
\hline
AA15-16  & RSV \emph{Aurora Australis}   & CL51    & AU       & 2015-10-22 & 2016-02-22 & 124 \\
HMNZSW16 & HMNZS \emph{Wellington}       & CHM 15k & NZ       & 2016-11-23 & 2016-12-19 & 27 \\
MARCUS   & RSV \emph{Aurora Australis}   & CT25K   & AU       & 2017-10-29 & 2018-03-26 & 149 \\
MICRE    & Macquarie Is. station         & CT25K   & AU/NZ    & 2016-04-03 & 2018-03-14 & 710 \\
NBP1704  & RV \emph{Nathaniel B. Palmer} & CHM 15k & NZ       & 2017-04-14 & 2017-06-08 & 55 \\
PS77/2   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2010-12-01 & 2011-02-04 & 65 \\
PS77/3   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2011-02-07 & 2011-04-14 & 66 \\
PS79/2   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2011-12-06 & 2012-01-02 & 27 \\
PS79/3   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2012-01-10 & 2012-03-10 & 61 \\
PS79/4   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2012-03-14 & 2012-04-08 & 26 \\
PS81/2   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2012-12-02 & 2013-01-18 & 47 \\
PS81/3   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-01-22 & 2013-03-17 & 55 \\
PS81/4   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-03-18 & 2013-04-16 & 30 \\
PS81/5   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-04-20 & 2013-05-23 & 33 \\
PS81/6   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-06-10 & 2013-08-12 & 63 \\
PS81/7   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-08-15 & 2013-10-14 & 60 \\
PS81/8   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-11-12 & 2013-12-14 & 31 \\
PS81/9   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-12-21 & 2014-03-02 & 71 \\
PS89     & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2014-12-05 & 2015-01-30 & 56 \\
PS96     & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2015-12-08 & 2016-02-14 & 68 \\
PS97     & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2016-02-15 & 2016-04-06 & 52 \\
PS103    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2016-12-18 & 2017-02-02 & 46 \\
PS104    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2017-02-08 & 2017-03-18 & 39 \\
PS111    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2018-01-21 & 2018-03-14 & 52 \\
PS112    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2018-03-18 & 2018-05-05 & 49 \\
PS117    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2018-12-18 & 2019-02-07 & 51 \\
PS118    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2019-02-18 & 2019-04-08 & 50 \\
PS123    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2021-01-10 & 2021-01-31 & 21 \\
PS124    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2021-02-03 & 2021-03-30 & 55 \\
TAN1502  & RV \emph{Tangaroa}            & CL51/31 & NZ       & 2015-01-20 & 2015-03-12 & 51 \\
TAN1702  & RV \emph{Tangaroa}            & CHM 15k & NZ       & 2017-03-09 & 2017-03-31 & 23 \\
TAN1802  & RV \emph{Tangaroa}            & CHM 15k & NZ       & 2018-02-07 & 2018-03-20 & 41 \\
\hline
\textbf{Total} &                         &         &          &            &            & \textbf{2421}\\
\end{tabular}
\normalsize
\end{table}

@@

The Vaisala CL51 (photo in Fig. \ref{fig:map}d) and CT25K are ceilometers
operating at a near-infrared wavelength of 910 nm and 905 nm, respectively.
CL51 can also be configured to emulate the Vaisala CL31. The maximum range is
15.4 km (CL51), 7.7 km (CL31 emulation mode with 5 m vertical resolution), and
7.5 km (CT25K). The vertical resolution is 10 m (5 m configurable) in CL51 and
30 m in CT25K. The sampling (temporal) resolution is configurable and in our
datasets it is about 6 s for CL51 on AA15‐16, 16 s for CT25K on MARCUS and
MICRE, 36 s for CL51 on RV \emph{Polarstern}, and about 2.37 s for CL51 with
CL31 emulation on TAN1502. The wavelength of 910 nm is affected by water vapour
absorption of about 20\% in the mid-latitudes \citep{wiegner2015,wiegner2019},
but we do not expect this to be a significant issue as explained in
\cite{kuma2021}.  The instrument data files containing raw uncalibrated
backscatter were first converted to Network Common Data Form (NetCDF) with
cl2nc (\url{https://github.com/peterkuma/cl2nc}) and then processed with the
ALCF (Section \ref{sec:alcf}) to produce absolutely calibrated attenuated
volume backscattering coefficinet (AVBC), cloud mask, cloud occurrence by
height and the total cloud fraction. Because the CT25K uses a very similar
wavelength to CL51, equivalent calculations as for CL51 were done assuming a
wavelength of 910 nm. The Vaisala CL51 and CT25K instruments were used on most
of the voyages and stations analysed here. Fig.  \ref{fig:example}a shows an
example of AVBC derived from the CL51 instrument data.

@@

The Lufft CHM 15k (photo in Fig. \ref{fig:map}c) is a ceilometer operating at a
near-infrared wavelength of 1064 nm. The maximum range is 15.4 km, the vertical
resolution is 5 m in the near range (up to 150 m) and 15 m above, the sampling
(temporal) resolution is 2 s, and the number of vertical levels is 1024.
NetCDF files containing uncalibrated backscatter produced by the instrument
were processed with the ALCF (Section \ref{sec:alcf}) to produce AVBC, cloud
mask, cloud occurrence by height and the total cloud fraction. The CHM 15k was
used on four voyages (HMNZSW16, TAN1702, TAN1802, and NBP1704).

@@

Absolute calibration of the observed backscatter was performed by comparing the
measured clear-sky molecular backscatter statistically with simulated clear-sky
molecular backscatter. AVBC was resampled to 5 min temporal resolution and 50 m
vertical resolution to increase signal-to-noise ratio, while having enough
resolution to detect small-scale cloud variability. Noise standard deviation
was calculated from AVBC at the highest range, where no clouds are expected.  A
cloud mask was calculated from AVBC using a fixed threshold of $\mathrm{2\times
10^{-6} m^{-1}sr^{-1}}$ after subtracting 5 standard deviations of range-scale
noise. Fig. \ref{fig:example}b shows an example of simulated Vaisala CL51
backscatter from ERA5 data, corresponding to a day of measurements by the
instrument on the PS81/3 voyage.

@@

Here, we use a free-running (i.e. \emph{not} nudged or using prescribed SST)
coupled GSRM simulation made for the purpse of climate projection.  nextGEMS
has so far produced four cycles of model runs. We used a Cycle 3 run
\emph{ngc3028} produced in 2023 \citep{nextgems2023a,nextgems2023b} for a model
time period of 20 January 2020 to 22 July 2025, of which we analysed the full
years 2021--2024 (inclusive). While a Cycle 4 run was available, we could not
use it due it to the unavailability of the necessary variables. The horizontal
resolution of ngc3028 is about 5 km.  The model output is available on 90
vertical levels and 3-hourly instantaneous temporal resolution.  Unlike current
general circulation models (GCMs), the strom-resolving version of ICON does not
use convective and cloud parameterisation but relies on explicit simulation of
convection and clouds on the model grid. While this makes the code development
simpler without having to rely on uncertain parameterisations, it can miss
smaller-scale clouds below the grid resolution.  Turbulence and cloud
microphysics are still parameterised in this model.

@@

ERA5 \citep{era5} is a reanalysis produced by the European Centre for
Medium-Range Weather Forecasts (ECMWF). It is based on a numerical weather
prediction model IFS version CY41R2.  The horizontal resolution is 0.25° in
latitude and longitude, which is about 28 km in the North--South direction and
14 km in the East--West direction at 60°S.  Internally, the model uses 137
vertical levels. Here, we use output at 1-hourly instantaeous time intervals,
except for radiation quantities, which are accumulations (from these we
calculate daily means).  Vertically-resolved quantities are made available on
37 pressure levels.

@@

\begin{figure}[b!]
\centering
\includegraphics[width=\textwidth]{img/ann.pdf}
\caption{
Artificial neural network (ANN) for prediction of precipitation in lidar
backscatter. \textbf{(a)} Diagram showing the TensorFlow structure of the ANN,
\textbf{(b)} randomly selected example samples of near-surface backscatter in
four categories (clear, fog, rain and snow), as determined by coincident
human-performed weather observations, \textbf{(c)} receiver operating
characteristic diagram of the ANN, \textbf{(d)} examples of 10-day time series
of human-observed (`HUM') and predicted precipitation based on an ANN trained
on all voyages (`ANN') and all voyages except for the shown voyage (`ANN2')
during three randomly selected voyages with the available data. Here, by
`randomly selected' we mean selected from the top of a permutation generated by
a pseudo-random number generator to prevent authors' bias in the selection.
}
\label{fig:ann}
\end{figure}

@@

We partitioned our data into two mutually-exclusive subsets by cyclonic
activity. For this purpose, we used a cyclone tracking algorithm to identify
extratropical and polar cyclones (ECs and PCs) over the SO in the reanalysis
and ICON data. We used the open source cyclone tracking package CyTRACK
\citep{perez-alarcon2024}.  Generally, what constitutes an EC is considered
relatively arbitrary due to the very large variability of ECs \citep{neu2013}.
We used the mean sea level pressure field and horizontal wind speed fields as
input to the CyTRACK algorithm. The algorithm uses pressure and wind speed
thresholds as well as tracking across time steps to identify cyclone centres
and radii. With this information, we could classify geographical areas as
either cyclonic or non-cyclonic. Due to a relatively small total area covered,
we chose a circle of a double radius (relative to one identified by CyTRACK)
centred at the cyclone centre as a cyclonic area for every time step and
cyclone. All other areas were identified as non-cyclonic. For identifying
cyclones in the observations and the reanalyses, ERA5 pressure and wind fields
were used as the input to CyTRACK.  This is justified by the fact that the
large-scale pressure and wind fields in ERA5 are likely sufficiently close to
reality. For identifying cyclones in ICON, its own pressure and wind fields
were used as the input to CyTRACK, because the model is free-running, and thus
the pressure and wind fields are different from reality.

>>>>

Increasing climate model resolution is a common way of improving model accuracy
\citep{mauritsen2022}. This is of course limited by the available computational
power and a tradeoff with increasing parametrisation complexity. Current
computational availability and acceleration from general-purpose computing on
graphics processing units (GPGPUs) is progressing to enable km-scale (also
called k-scale) Earth system models (ESMs) and coupled atmosphere--ocean
general circulation models (AOGCMs) in research conditions today and
operationally in the forthcoming years. Therefore, it represents a natural
advancement in climate modeling.  Global strom-resolving models (GSRMs) are
emerging as a new front in the development of high-resolution global climate
models, with horizontal grid resolutions of about 2--8 km
\citep{satoh2019,stevens2019}. This is enough to resolve mesoscale convective
storms, but smaller-scale convective plumes and smaller-scale cloud structure
remain unresolved. At a 5-km scale, non-hydrostatic processes also become
important, and for this reason such models are generally non-hydrostatic. The
terms global cloud-resolving models or global convection-permitting/-resolving
models are also sometimes used interchangeably with GSRMs but imply that clouds
or convection are resolved explicitly, which is not entirely true for GSRMs, as
this would require an even higher horizontal resolution \citep{satoh2019}.
Representative of these efforts is the DYnamics of the Atmospheric general
circulation Modeled On Non-hydrostatic Domains (DYAMOND) project
\citep{stevens2019,dyamond}, which is an intercomparison of nine global
storm-resolving models (GSRMs) on two 40-day time periods in summer (1 August
-- 10 September 2016) and winter (20 January -- 1 March 2020). A new one-year
GSRM intercomparison is currently proposed by \cite{takasuka2024}, with the
hope of also evaluating the seasonal cycle and large-scale circulation.  An
alternative to using a computationally costly GSRM is to train an artificial
neural network on GSRM output and use it for subgrid-scale clouds, as done with
the GSRM ICON by \cite{grundner2022} and \cite{grundner2023}.

@@

The Southern Ocean (SO) is known to be a problematic region for climate model
biases due to reasons such as a lack of surface and in situ observations and
being a lower priority region for numerical weather prediction (NWP) and
climate model development because it is far away from populated areas of
interest.  Nevertheless, radiation biases and changes over an area of its size
have a very substantial influence on the global climate, and the SO is an
important part of the global ocean conveyor belt.  Marine clouds have
a disproportionate effect on top of atmosphere (TOA) SW radiation due to the
relatively low albedo of the sea surface.  The relative longitudinal symmetry
of the SO means that model cloud biases tend to be similar across longitudes.
Here, we conventionally refer to the SO as ocean regions south of 40°S,
low-latitude SO as 40--55°S and high-latitude SO as south of 55°S.

@@

The main aim of this study is to evaluate the GSRM version of ICON, developed
jointly by nextGEMS, Deutscher Wetterdienst, Max-Planck-Institute for
Meteorology, Deutsches Klimarechenzentrum (DKRZ), Karlsruhe Institute of
Technology, and the Center for Climate Systems Modeling.  Previous studies have
identified substantial large-scale biases in climate model clouds over the SO,
affecting sea surface temperature and the Earth’s albedo. Our aim is to
quantify how well the GSRM ICON is simulating clouds in this region,
particularly in light of the fact that subgrid-scale clouds and convection are
not parameterized in this model. This region is mostly dominated by boundary
layer clouds generated by shallow convection, and these are problematic to
observe by spaceborne lidars and radars, which are affected by attenuation by
overlapping and thick clouds and ground clutter, respectively.  Specifically,
the radar on CloudSat and lidar on CALIPSO (neither of them operational any
more) are affected by the abovementioned issues, resulting in a strong
underestimation of cloud occurrence below 2 km relative to ground-based lidar
observations \citep{mcerlich2021}.  This, in turn, can lead to systematic
biases in low clouds in climate models, which are frequently evaluated against
CloudSat--CALIPSO products. Reanalyses can also suffer from cloud biases, as
these are usually parametrised in their atmospheric component, and also in
regions where input observations are sparse.  This makes them a problematic
reference for clouds over the SO, and any biases relative to a reanalysis
should be interpreted with caution. Instead, we chose to use a large set of
ship-based observations conducted with ceilometers and lidars on board of the
RV \emph{Polarstern} and other voyages and stations as a reference for the
model evaluation.

Altogether, we analysed about 2400 days of data from 31 voyages, and one
sub-Antarctic station covering diverse longitudes and latitudes of the SO. To
achieve a like-for-like comparison with the model, we used a ground-based lidar
simulator called the Automatic Lidar and Ceilometer Framework [ALCF;
\cite{kuma2021}]. We contrasted the results with the European Centre for
Medium-Range Weather Forecasts (ECMWF) Reanalysis 5 [ERA5; \cite{era5}] and the
Modern-Era Retrospective analysis for Research and Applications, Version 2
[MERRA-2; \cite{gelaro2017}].

@@

Together, we analysed data from 31 voyages of RV \emph{Polarstern}, resupply
vessel (RSV) \emph{Aurora Australis}, RV \emph{Tangaroa}, RV \emph{Nathaniel B.
Palmer}, Her Majesty's New Zealand Ship (HMNZS) \emph{Wellington} and one
sub-Antarctic station (Macquarie Island) in the SO south of 40°S between 2010
and 2021. Fig. \ref{fig:map} shows a map of the voyages, Table
\ref{tab:voyages} list the voyages, campaigns, and stations, and Table
\ref{tab:voyage-references} lists references where available. Altogether, the
voyages and station dataset comprised 2421 days of data south of 40°S, but the
availability of ceilometer data was slightly smaller due to gaps in
measurements.

@@

The data sources contained ceilometer observations captured by the Vaisala CL51
 perating at a wavelenght of 910 nm, the Vaisala CT25K operating at 905 nm, and
the Lufft CHM 15k operating at 1064 nm, described in detail below (Sections
\ref{sec:cl51} and \ref{sec:chm15k}). A ceilometer is a low-power near-infrared
vertically pointing lidar with the purpose of measuring the cloud base, but it
also measures the full vertical structure of clouds as long as the laser signal
is not attenuated by thick clouds, which can be used to infer additional
information such as a cloud mask and cloud occurrence by height.

@@

\begin{table}[p!]
% TODO: Add sampling intervals.
\caption{
An overview of the analysed voyages, campaigns, and stations. Start, end, and
the number of days (UTC; inclusive) refer to the time period when the vessel
was south of 40°S.  Abbreviations: ceilometer (ceil.), Australia (AU), New
Zealand (NZ), South America (SA), Atlantic Ocean (AO), and Africa (AF). The
number of days is rounded to the nearest integer. CL51/31 indicates CL51
configured to emulate CL31.
}
\label{tab:voyages}
\small
\begin{tabular}{llllllr}
\textbf{Name} & \textbf{Vessel or station} & \textbf{Ceil.} & \textbf{Region} & \textbf{Start} & \textbf{End} & \textbf{Days}\\
\hline
AA15-16  & RSV \emph{Aurora Australis}   & CL51    & AU       & 2015-10-22 & 2016-02-22 & 124 \\
HMNZSW16 & HMNZS \emph{Wellington}       & CHM 15k & NZ       & 2016-11-23 & 2016-12-19 & 27 \\
MARCUS   & RSV \emph{Aurora Australis}   & CT25K   & AU       & 2017-10-29 & 2018-03-26 & 149 \\
MICRE    & Macquarie Is. station         & CT25K   & AU/NZ    & 2016-04-03 & 2018-03-14 & 710 \\
NBP1704  & RV \emph{Nathaniel B. Palmer} & CHM 15k & NZ       & 2017-04-14 & 2017-06-08 & 55 \\
PS77/2   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2010-12-01 & 2011-02-04 & 65 \\
PS77/3   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2011-02-07 & 2011-04-14 & 66 \\
PS79/2   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2011-12-06 & 2012-01-02 & 27 \\
PS79/3   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2012-01-10 & 2012-03-10 & 61 \\
PS79/4   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2012-03-14 & 2012-04-08 & 26 \\
PS81/2   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2012-12-02 & 2013-01-18 & 47 \\
PS81/3   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-01-22 & 2013-03-17 & 55 \\
PS81/4   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-03-18 & 2013-04-16 & 30 \\
PS81/5   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-04-20 & 2013-05-23 & 33 \\
PS81/6   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-06-10 & 2013-08-12 & 63 \\
PS81/7   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-08-15 & 2013-10-14 & 60 \\
PS81/8   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-11-12 & 2013-12-14 & 31 \\
PS81/9   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-12-21 & 2014-03-02 & 71 \\
PS89     & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2014-12-05 & 2015-01-30 & 56 \\
PS96     & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2015-12-08 & 2016-02-14 & 68 \\
PS97     & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2016-02-15 & 2016-04-06 & 52 \\
PS103    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2016-12-18 & 2017-02-02 & 46 \\
PS104    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2017-02-08 & 2017-03-18 & 39 \\
PS111    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2018-01-21 & 2018-03-14 & 52 \\
PS112    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2018-03-18 & 2018-05-05 & 49 \\
PS117    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2018-12-18 & 2019-02-07 & 51 \\
PS118    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2019-02-18 & 2019-04-08 & 50 \\
PS123    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2021-01-10 & 2021-01-31 & 21 \\
PS124    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2021-02-03 & 2021-03-30 & 55 \\
TAN1502  & RV \emph{Tangaroa}            & CL51/31 & NZ       & 2015-01-20 & 2015-03-12 & 51 \\
TAN1702  & RV \emph{Tangaroa}            & CHM 15k & NZ       & 2017-03-09 & 2017-03-31 & 23 \\
TAN1802  & RV \emph{Tangaroa}            & CHM 15k & NZ       & 2018-02-07 & 2018-03-20 & 41 \\
\hline
\textbf{Total} &                         &         &          &            &            & \textbf{2421}\\
\end{tabular}
\normalsize
\end{table}

@@

The Vaisala CL51 (photo in Fig. \ref{fig:map}d) and CT25K are ceilometers
operating at a near-infrared wavelength of 910 nm and 905 nm, respectively.
CL51 can also be configured to emulate the Vaisala CL31. The maximum range is
15.4 km (CL51), 7.7 km (CL31 emulation mode with 5 m vertical resolution), and
7.5 km (CT25K). The vertical resolution is 10 m (5 m configurable) in CL51 and
30 m in CT25K. The sampling (temporal) resolution is configurable, and in our
datasets it is about 6 s for CL51 on AA15‐16, 16 s for CT25K on MARCUS and
MICRE, 36 s for CL51 on RV \emph{Polarstern}, and about 2.37 s for CL51 with
CL31 emulation on TAN1502. The wavelength of 910 nm is affected by water vapour
absorption of about 20\% in the mid-latitudes \citep{wiegner2015,wiegner2019},
but we do not expect this to be a significant issue as explained in
\cite{kuma2021}.  The instrument data files containing raw uncalibrated
backscatter were first converted to Network Common Data Form (NetCDF) with
cl2nc (\url{https://github.com/peterkuma/cl2nc}) and then processed with the
ALCF (Section \ref{sec:alcf}) to produce absolutely calibrated attenuated
volume backscattering coefficient (AVBC), cloud mask, cloud occurrence by
height, and the total cloud fraction. Because the CT25K uses a very similar
wavelength to CL51, equivalent calculations as for CL51 were done assuming a
wavelength of 910 nm. The Vaisala CL51 and CT25K instruments were used on most
of the voyages and stations analysed here. Fig.  \ref{fig:example}a shows an
example of AVBC derived from the CL51 instrument data.

@@

The Lufft CHM 15k (photo in Fig. \ref{fig:map}c) is a ceilometer operating at a
near-infrared wavelength of 1064 nm. The maximum range is 15.4 km, the vertical
resolution is 5 m in the near range (up to 150 m) and 15 m above, the sampling
(temporal) resolution is 2 s, and the number of vertical levels is 1024.
NetCDF files containing uncalibrated backscatter produced by the instrument
were processed with the ALCF (Section \ref{sec:alcf}) to produce AVBC, cloud
mask, cloud occurrence by height, and the total cloud fraction. The CHM 15k was
used on four voyages (HMNZSW16, TAN1702, TAN1802, and NBP1704).

@@

Absolute calibration of the observed backscatter was performed by comparing the
measured clear-sky molecular backscatter statistically with simulated clear-sky
molecular backscatter. AVBC was resampled to 5 min temporal resolution and 50 m
vertical resolution to increase signal-to-noise ratio while having enough
resolution to detect small-scale cloud variability. The noise standard
deviation was calculated from AVBC at the highest range, where no clouds are
expected.  A cloud mask was calculated from AVBC using a fixed threshold of
$\mathrm{2\times 10^{-6} m^{-1}sr^{-1}}$ after subtracting 5 standard
deviations of range-scale noise. Fig. \ref{fig:example}b shows an example of
simulated Vaisala CL51 backscatter from ERA5 data, corresponding to a day of
measurements by the instrument on the PS81/3 voyage.

@@

Here, we use a free-running (i.e., \emph{not} nudged or using prescribed SST)
coupled GSRM simulation made for the purpose of climate projection.  nextGEMS
has so far produced four cycles of model runs. We used a Cycle 3 run
\emph{ngc3028} produced in 2023 \citep{nextgems2023a,nextgems2023b} for a model
time period of 20 January 2020 to 22 July 2025, of which we analysed the full
years 2021--2024 (inclusive). While a Cycle 4 run was available, we could not
use it due to the unavailability of the necessary variables. The horizontal
resolution of ngc3028 is about 5 km.  The model output is available on 90
vertical levels and 3-hourly instantaneous temporal resolution.  Unlike current
general circulation models (GCMs), the strom-resolving version of ICON does not
use convective and cloud parameterization but relies on explicit simulation of
convection and clouds on the model grid. While this makes the code development
simpler without having to rely on uncertain parameterizations, it can miss
smaller-scale clouds below the grid resolution.  Turbulence and cloud
microphysics are still parameterized in this model.

@@

ERA5 \citep{era5} is a reanalysis produced by the European Centre for
Medium-Range Weather Forecasts (ECMWF). It is based on a numerical weather
prediction model IFS version CY41R2.  The horizontal resolution is 0.25° in
latitude and longitude, which is about 28 km in the North--South direction and
14 km in the East--West direction at 60°S.  Internally, the model uses 137
vertical levels. Here, we use output at 1-hourly instantaeous time intervals,
except for radiation quantities, which are accumulations (from these we
calculate daily means).  Vertically resolved quantities are made available on
37 pressure levels.

@@

\begin{figure}[b!]
\centering
\includegraphics[width=\textwidth]{img/ann.pdf}
\caption{
Artificial neural network (ANN) for prediction of precipitation in lidar
backscatter. \textbf{(a)} Diagram showing the TensorFlow structure of the ANN,
\textbf{(b)} randomly selected example samples of near-surface backscatter in
four categories (clear, fog, rain, and snow), as determined by coincident
human-performed weather observations, \textbf{(c)} receiver operating
characteristic diagram of the ANN, \textbf{(d)} examples of 10-day time series
of human-observed (`HUM') and predicted precipitation based on an ANN trained
on all voyages (`ANN') and all voyages except for the shown voyage (`ANN2')
during three randomly selected voyages with the available data. Here, by
`randomly selected' we mean selected from the top of a permutation generated by
a pseudo-random number generator to prevent authors' bias in the selection.
}
\label{fig:ann}
\end{figure}

@@

We partitioned our data into two mutually exclusive subsets by cyclonic
activity. For this purpose, we used a cyclone tracking algorithm to identify
extratropical and polar cyclones (ECs and PCs) over the SO in the reanalysis
and ICON data. We used the open source cyclone tracking package CyTRACK
\citep{perez-alarcon2024}.  Generally, what constitutes an EC is considered
relatively arbitrary due to the very large variability of ECs \citep{neu2013}.
We used the mean sea level pressure field and horizontal wind speed fields as
input to the CyTRACK algorithm. The algorithm uses pressure and wind speed
thresholds as well as tracking across time steps to identify cyclone centres
and radii. With this information, we could classify geographical areas as
either cyclonic or non-cyclonic. Due to a relatively small total area covered,
we chose a circle of a double radius (relative to one identified by CyTRACK)
centred at the cyclone centre as a cyclonic area for every time step and
cyclone. All other areas were identified as non-cyclonic. For identifying
cyclones in the observations and the reanalyses, ERA5 pressure and wind fields
were used as the input to CyTRACK.  This is justified by the fact that the
large-scale pressure and wind fields in ERA5 are likely sufficiently close to
reality. For identifying cyclones in ICON, its own pressure and wind fields
were used as the input to CyTRACK, because the model is free-running, and thus
the pressure and wind fields are different from reality.
